### 2014-03-24 doc/paper.tex
Import paper into LaTeX

\documentclass[conference]{acmsiggraph}

\usepackage{mdwlist}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{enumitem}

%%%%%%%%%%%%%%%
% Information %
%%%%%%%%%%%%%%%

\title{Plant Modeling Using Environmental Parameters}

\author{
  Andre Philipp\thanks{e-mail:philipp.andy@gmail.com}
  \qquad
  Andy Spencer\thanks{e-mail:e-mail:andy753421@ucla.edu}
  \qquad
  Haakon Garseg Mørk\thanks{e-mail:haakongmork@gmail.com}
  \\
  University of California, Los Angeles \\
  CS275 Artificial Life for Computer Graphics and Vision
}

\pdfauthor{Andre Philipp, Andy Spencer, Haakon Garseg Mørk}

\keywords{radiosity, global illumination, constant time}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%
% Title Page %
%%%%%%%%%%%%%%

\maketitle

\begin{abstract}

The project will focus on modeling natural looking plants using simulated
environment parameters to control growth and evolution of different plant
species. Genetic algorithms or other evolutionary techniques will be used to
evolve plant suitable for growth in various environments.

Potential environmental conditions:

\begin{itemize*}
  \item Availability of water
  \item Amount of sunlight
  \item Soil conditions
  \item Growth elevation
\end{itemize*}

Potential plant species parameters:

\begin{itemize*}
  \item Rate of water loss
  \item Food and water storage capability
  \item Total mass/volume
  \item Crown height
  \item Leaf area
  \item Root structures
\end{itemize*}

\end{abstract}

\keywordlist

%%%%%%%%%%%%
% Content %
%%%%%%%%%%%%

\section{Introduction}

Karl Sims introduced Panspermia in 1990.

LSystem

We closely examined different existing frameworks that we considered using for
our project. The two frameworks that were the best candidates was
OpenAlea\cite{openalea} and Algorithmic Botany\cite{abotany}.

\section{Architecture and Models}

\includegraphics[width=\columnwidth]{images/architecture.png}

\section{Simulation}

The simulation of the plant environment is governed by ticks. For each tick that
goes, all plant objects becomes older. We wanted to have a realistic approach to
have plant grow in the real world, and therefore the first 4 ticks, the plant is
growing as a seed under ground. On the fifth tick it starts to grow above
ground. When the seventh tick is reached, the plant becomes mature and grows
flowers and leaves, if the gene is capable of it. The plant object then spreads
it seeds within a certain radius. When the season is over, the plant dies, but
it’s children grows up with either the same gene, or and mutation of the gene.

\section{Genetics}

In plant biology it’s defined two types of reproduction. Asexual reproduction
and sexual reproduction. Asexual reproduction can either be a clone or a
mutation of its parent. Sexual reproduction creates a new offspring by combining
the genetics from its parents.\cite{plantrepo}

\section{Conclusion}

Gene:

\begin{description}[leftmargin=!,labelindent=0.2in,labelwidth=0.1in]
  \item[F]   grow stem
  \item[L]   grow flower
  \item[X]   does not correspond to any drawing action and is used to control
             the evolution of the curve.\cite{lsystems}
  \item[n, e, s, w] \hfill \\
             bend plants growing direction either north, east, south or west.
             The angle is dependent on the genes angle.
  \item[{[}] push matrix
  \item[{]}] pop matrix
\end{description}

Example of a plant at first generation:

\begin{description}[leftmargin=!,labelindent=0.2in,labelwidth=0.4in]
  \item[Angle] 15 degrees
  \item[Axiom] X
  \item[Rules] \{ ``Fs[[F]eX]nF[wFX]F'' , ``FF'' \}
\end{description}

Notice that this plant will not start with flowers (`L').

Generally when the plant grows, it takes a random character from the LString
(but not `[' or `]') and substitutes this character with one random rule from
the plant gene. At the seventh tick of the season, the plant is mature and will
clone itself. There's two ways that this can happen. Either the new plant will
grow up with the same gene, or the gene will mutate and the new plant will be
given the mutated gene. The mutation happens in the set of the gene rules. A
random rule is picked out to be mutated. Again there is two possible outcomes.
Either a character in the rule will be deleted, or a random string from a
predefined mutation array will be added at a random position in the rule. We
want to simulate subtle changes in the gene as the mutation isn't suppose to be
extreme in any ways. Therefore the mutation array can be in the following way \{
``F'', ``L'', ``xF'', ``xL'', ``xX'', ``[X]'', ``[F]'', ``[L]'', ``[xF]'',
``[xL]'' \}. `x' is substituted with a random direction to grow, i.e. `n', `e',
`s' or `w'.

\section{Environmental parameters}

There are three basic environmental parameters which has an effect on the
plants. The amount of sun (0-100\%), water (0-100\%) and the nutrition available
in the ground.

\section{Graphics}

\section{Conclusion}

%%%%%%%%%%%%%%
% References %
%%%%%%%%%%%%%%

\bibliographystyle{acmsiggraph}
\bibliography{paper}

\end{document}

### 2014-03-24 src/plant_project/Drawer.java
Position camera a little higher

@@ -458,6 +458,7 @@ public class Drawer extends WindowAdapter implements KeyListener,
 		gl.glLoadIdentity();
 
 		// Camera 1
+		gl.glTranslated(0f, -1f, 0f);
 		gl.glRotated( this.rotation[0], 1f, 0f, 0f); // tiltx
 		gl.glRotated( this.rotation[2], 0f, 0f, 1f); // tiltz
 

### 2014-03-24 src/plant_project/Main.java
Add test plants

@@ -121,6 +121,20 @@ public class Main {
 		this.plants = new ArrayList<Plant>();
 		this.drawer = new Drawer(this);
 
+		/*
+		plants.add(new Plant(0, 0, new Gene(15,
+			new String[] { "Fn[[X]sXL]eF[sFX]X", "FF" },
+			new Color(0.0f, 0.8f, 0.2f),
+			new Color(0.3f, 0.3f, 0.8f),
+			2, Plant.MAX_ITERATION)));
+
+		plants.add(new Plant(1, 0, new Gene(15,
+			new String[] { "Fs[[FL]eX]nF[wFXL]F", "FF" },
+			new Color(0.2f, 0.9f, 0.9f),
+			new Color(0.8f, 0.3f, 0.3f),
+			2, Plant.MAX_ITERATION)));
+		*/
+
 		System.out.format("%-16s %3d%% (s=?)\n", "Sun:", sun);
 		System.out.format("%-16s %3d%% (w=?)\n", "Water:", water);
 		System.out.format("%-16s %3d%% (m=?)\n", "Mutation Prob:", mutation);

### 2014-03-24 src/plant_project/Web.java
Play around with Java Applets

@@ -1,0 +1,10 @@
+package plant_project;
+
+import java.applet.Applet;
+import java.awt.Graphics;
+
+public class Web extends Applet {
+	public void paint(Graphics g) {
+		g.drawString("Hello world!", 50, 25);
+	}
+}

### 2014-03-24 src/plant_project/Drawer.java
Move rules into a separate LRule class
- This introduces probabilities and productions similar to what is
  described in the "Stochastic grammars" grammars section of wikipedia.

- Also move LString definitions to LSystems class instead of hard-coding
  in Drawer.

@@ -32,16 +32,6 @@ public class Drawer extends WindowAdapter implements KeyListener,
 		MouseListener, MouseWheelListener, MouseMotionListener,
 		GLEventListener {
 
-	private final static char NOTHING     = 'X';
-	private final static char GROW        = 'F';
-	private final static char GROW_FlOWER = 'L';
-	private final static char NORTH       = 'n';
-	private final static char EAST        = 'e';
-	private final static char WEST        = 'w';
-	private final static char SOUTH       = 's';
-	private final static char PUSH        = '[';
-	private final static char POP         = ']';
-
 	public final static float NORTH_BOUND =  4.0f;
 	public final static float EAST_BOUND  =  4.0f;
 	public final static float WEST_BOUND  = -4.0f;
@@ -182,58 +172,82 @@ public class Drawer extends WindowAdapter implements KeyListener,
 		// move to plant origin
 		gl.glTranslatef(plant.getX(), 0.0f, plant.getY());
 		
-		int fCount = CharMatcher.is('F').countIn(lString);
+		int fCount = CharMatcher.is(LSystem.GROW).countIn(lString);
 		float step 		= 0.3f;
 		float stemLevel = 1.0f;
 		for (int i = 0; i < lString.length(); i++) {
 			char c = lString.charAt(i);
 			switch (c) {
-			case NOTHING:
+			case LSystem.NOTHING:
 				// do nothing
 				break;
-			case GROW:
+			case LSystem.GROW:
 				// drawLine
-				float lineLength 	= 0.1f;
-				float radius 		=  stemLevel * (fCount/(float)Math.sqrt(fCount))/2000;
+				float lineLength = 0.1f;
+				float radius     = stemLevel * (fCount/(float)Math.sqrt(fCount))/2000;
 				this.drawStem(gl, stemColor, lineLength, radius);
 				gl.glTranslatef(0, lineLength, 0.0f);
 				break;
-			case GROW_FlOWER:
+			case LSystem.FLOWER:
 				if (plant.getAge() >= Plant.MAX_ITERATION)
 					this.drawFlower(gl, flowerColor);
 				break;
-			case PUSH:
+			case LSystem.PUSH:
 				// push position and angle
 				if (stemLevel - step >= 0.1f)
[…]

### 2014-03-24 src/plant_project/Gene.java

@@ -3,11 +3,16 @@ package plant_project;
 import java.awt.Color;
 
 public class Gene {
-	private static final char[] ORIENTATIONS = new char[] { 'n', 's', 'e', 'w' };
-	private static final String[] MUTATIONS = new String[] { "F", "L", "xF",
-		"xL", "xX", "[X]", "[F]", "[L]", "[xF]", "[xL]" };
+	private static final char[] ORIENTATIONS = new char[] {
+		'n', 's', 'e', 'w'
+	};
+	private static final String[] MUTATIONS = new String[] {
+		"G", "xG", "[G]", "[xG]",
+		"F", "xF", "[F]", "[xF]",
+		"[X]", "xX",
+	};
 	private int angle;
-	private String[] rules;
+	private LRule[] rules;
 	private Color stemColor;
 	private Color flowerColor;
 
@@ -18,7 +23,7 @@ public class Gene {
 		return angle;
 	}
 
-	public String[] getRules() {
+	public LRule[] getRules() {
 		return rules;
 	}
 
@@ -38,7 +43,7 @@ public class Gene {
 		return maturityAge;
 	}
 
-	public Gene(int angle, String[] rules, Color stemColor, Color flowerColor,
+	public Gene(int angle, LRule[] rules, Color stemColor, Color flowerColor,
 			int reproductionCount, int maturityAge) {
 		super();
 		this.angle = angle;
@@ -58,12 +63,12 @@ public class Gene {
 		int i = (int) (Math.random() * 100);
 		// MUTATION
 		if (i > 0 && i <= Main.mutation) {
-			String[] newRules = rules.clone();
+			LRule[] newRules = rules.clone();
 			// pick random rule to mutate
 			int mutationRuleIndex = (int) (Math.random() * (newRules.length - 1));
 			newRules[mutationRuleIndex] = mutate(newRules[mutationRuleIndex]);
-			System.out.println(rules[mutationRuleIndex]);
-			System.out.println(newRules[mutationRuleIndex]);
+			System.out.println("old: " + rules[mutationRuleIndex]);
+			System.out.println("new: " + newRules[mutationRuleIndex]);
 			Color stemColor = new Color((float) Math.random(),
 					(float) Math.random(), (float) Math.random());
 			Color flowerColor = new Color((float) Math.random(),
@@ -80,8 +85,9 @@ public class Gene {
 	public static final int ADD = 1;
 	public static final int DELETE = 0;
 
[…]

### 2014-03-24 src/plant_project/LRule.java

@@ -1,0 +1,27 @@
+package plant_project;
+
+public class LRule {
+	public double prob;
+	public char   find;
+	public String replace;
+
+	public LRule(String str) {
+		String[] parts = str.split(":");
+
+		this.prob    = Double.parseDouble(parts[0]);
+		this.find    = parts[1].charAt(0);
+		this.replace = parts[2];
+	}
+
+	public LRule(double probability, char find, String replace) {
+		this.prob    = probability;
+		this.find    = find;
+		this.replace = replace;
+	}
+
+	public String toString() {
+		return String.format("[%4.2f] %s -> %s",
+			this.prob, this.find, this.replace);
+	}
+}
+

### 2014-03-24 src/plant_project/LSystem.java

@@ -1,16 +1,39 @@
 package plant_project;
 
-public class LSystem {
-	public final static char VARIABLES[] = {'X', 'F'};
-	public final static char FLOWER = 'L';
-	public final static char AXIOM = 'X';
-	public final static char POSITIVE = '+';
-	public final static char NEGATIVE = '-';
-	
-	public static String iterate(String lString, String[] rules){
-		char randomVariable = VARIABLES[(int)(Math.random()*VARIABLES.length)];
-		lString = lString.replace(Character.toString(randomVariable), rules[(int)Math.random()*rules.length]);
-		return lString;
-	}
+import java.util.Random;
 
+public class LSystem {
+	public final static char AXIOM       = 'X';
+
+	public final static char NOTHING     = 'X';
+	public final static char GROW        = 'G';
+	public final static char LEAF        = 'L';
+	public final static char FLOWER      = 'F';
+	public final static char NORTH       = 'n';
+	public final static char EAST        = 'e';
+	public final static char WEST        = 'w';
+	public final static char SOUTH       = 's';
+	public final static char PUSH        = '[';
+	public final static char POP         = ']';
+
+	public static String iterate(String lString, LRule[] rules) {
+		StringBuilder output = new StringBuilder();
+
+		for (int i = 0; i < lString.length(); i++) {
+			boolean found = false;
+			char chr = lString.charAt(i);
+			for (LRule rule : rules) {
+				if (chr == rule.find && Math.random() < rule.prob) {
+					output.append(rule.replace);
+					found = true;
+					break;
+				}
+			}
+			if (!found) {
+				output.append(chr);
+			}
+		}
+
+		return output.toString();
+	}
 }

### 2014-03-24 src/plant_project/Main.java

@@ -42,19 +42,26 @@ public class Main {
 	public void keyPress(char c, String in)
 	{
 		switch (c) {
-			case 'p':// plant p=(rules seperated by ,), reproducecount,
-				// maturityage
+			case 'p':
+				// plant p=(probability:find:replace; ...), reproducecount, maturityage
 				if (in.length() < 3) {
-					plants.add(new Plant(1, 0, new Gene(15, new String[] {
-						"Fs[[F]eX]nF[wFX]F", "FF" }, new Color(0.2f, 0.9f,
-							0.0f), Color.BLUE, 2, Plant.MAX_ITERATION)));
+					plants.add(new Plant(1, 0, new Gene(15,
+						new LRule[] { new LRule(0.5, 'X', "Gs[[G]eX]nG[wGX]G"),
+						              new LRule(0.5, 'G', "GG") },
+						new Color(0.2f, 0.9f, 0.0f),
+						Color.BLUE,
+						2, Plant.MAX_ITERATION)));
 				} else {
 					String[] plantElements = in.split("=")[1].split(";");
-					String[] rules = plantElements[0].substring(1,
-							plantElements[0].length() - 2).split(",");
-					plants.add(new Plant(0, 0, new Gene(15, rules, Color.GREEN,
-									Color.BLUE, Integer.parseInt(plantElements[1]),
-									Integer.parseInt(plantElements[2]))));
+					String[] ruleStrings = plantElements[0].substring(1,
+							plantElements[0].length()-2).split(",");
+					LRule rules[] = new LRule[ruleStrings.length];
+					for (int i = 0; i < ruleStrings.length; i++)
+						rules[i] = new LRule(ruleStrings[i]);
+					plants.add(new Plant(0, 0, new Gene(15,
+						rules, Color.GREEN, Color.BLUE,
+						Integer.parseInt(plantElements[1]),
+						Integer.parseInt(plantElements[2]))));
 				}
 				System.out.println("planted in 0, 0!");
 				break;
@@ -121,19 +128,27 @@ public class Main {
 		this.plants = new ArrayList<Plant>();
 		this.drawer = new Drawer(this);
 
-		/*
[…]

### 2014-03-24 src/plant_project/Drawer.java
Add leaves

@@ -111,6 +111,23 @@ public class Drawer extends WindowAdapter implements KeyListener,
 		gl.glEnd();
 	}
 
+	private void drawLeaf(GL2 gl, Color color, float size) {
+		// Draw stemp vertex
+		gl.glColor3ub((byte)color.getRed(),
+			      (byte)color.getGreen(),
+			      (byte)color.getBlue());
+		gl.glBegin(GL2.GL_TRIANGLE_FAN);
+
+		gl.glNormal3f( 0.0f, 0.5f, 0.0f); gl.glVertex3f(0.0f, 0.0f, 0.0f);
+		gl.glNormal3f( 0.3f, 0.0f, 0.0f); gl.glVertex3f(-0.5f*size, 0.2f*size,  0.5f*size);
+		gl.glNormal3f( 0.2f, 0.1f, 0.0f); gl.glVertex3f(-0.4f*size, 0.1f*size,  0.8f*size);
+		gl.glNormal3f( 0.0f, 0.5f, 0.0f); gl.glVertex3f( 0.0f*size, 0.0f*size,  1.0f*size);
+		gl.glNormal3f(-0.3f, 0.1f, 0.0f); gl.glVertex3f( 0.4f*size, 0.1f*size,  0.8f*size);
+		gl.glNormal3f(-0.3f, 0.0f, 0.0f); gl.glVertex3f( 0.5f*size, 0.2f*size,  0.5f*size);
+
+		gl.glEnd();
+	}
+
 	private void drawPedals(GL2 gl, Color color,
 			int pedals, float length, float radius, float norm) {
 
@@ -167,13 +184,15 @@ public class Drawer extends WindowAdapter implements KeyListener,
 
 		Color stemColor   = plant.getGene().getStemColor();
 		Color flowerColor = plant.getGene().getFlowerColor();
+		float stemWidth   = plant.getGene().getStemWidth();
+		float leafSize    = plant.getGene().getLeafSize();
 
 		gl.glPushMatrix();
 		// move to plant origin
 		gl.glTranslatef(plant.getX(), 0.0f, plant.getY());
 		
 		int fCount = CharMatcher.is(LSystem.GROW).countIn(lString);
-		float step 		= 0.3f;
+		float step      = 0.3f;
 		float stemLevel = 1.0f;
 		for (int i = 0; i < lString.length(); i++) {
 			char c = lString.charAt(i);
@@ -184,9 +203,13 @@ public class Drawer extends WindowAdapter implements KeyListener,
[…]

### 2014-03-24 src/plant_project/Gene.java

@@ -19,6 +19,9 @@ public class Gene {
 	private int reproductionCount;
 	private int maturityAge;
 
+	private float stemWidth;
+	private float leafSize;
+
 	public int getAngle() {
 		return angle;
 	}
@@ -43,8 +46,16 @@ public class Gene {
 		return maturityAge;
 	}
 
+	public float getStemWidth() {
+		return stemWidth;
+	}
+
+	public float getLeafSize() {
+		return leafSize;
+	}
+
 	public Gene(int angle, LRule[] rules, Color stemColor, Color flowerColor,
-			int reproductionCount, int maturityAge) {
+			int reproductionCount, int maturityAge, float stemWidth, float leafSize) {
 		super();
 		this.angle = angle;
 		this.rules = rules;
@@ -52,6 +63,8 @@ public class Gene {
 		this.flowerColor = flowerColor;
 		this.reproductionCount = reproductionCount;
 		this.maturityAge = maturityAge;
+		this.stemWidth = stemWidth;
+		this.leafSize = leafSize;
 	}
 
 	/**
@@ -74,11 +87,13 @@ public class Gene {
 			Color flowerColor = new Color((float) Math.random(),
 					(float) Math.random(), (float) Math.random());
 			return new Gene(angle, newRules, stemColor, flowerColor,
-					reproductionCount, maturityAge);
+					reproductionCount, maturityAge,
+					stemWidth, leafSize);
 		} else {
 			// NO MUTATION
 			return new Gene(angle, rules, stemColor, flowerColor,
-					reproductionCount, maturityAge);
+					reproductionCount, maturityAge,
+					stemWidth, leafSize);
 		}
 	}
 

### 2014-03-24 src/plant_project/Main.java

@@ -50,7 +50,8 @@ public class Main {
 						              new LRule(0.5, 'G', "GG") },
 						new Color(0.2f, 0.9f, 0.0f),
 						Color.BLUE,
-						2, Plant.MAX_ITERATION)));
+						2, Plant.MAX_ITERATION,
+						0.01f, 0.1f)));
 				} else {
 					String[] plantElements = in.split("=")[1].split(";");
 					String[] ruleStrings = plantElements[0].substring(1,
@@ -61,7 +62,8 @@ public class Main {
 					plants.add(new Plant(0, 0, new Gene(15,
 						rules, Color.GREEN, Color.BLUE,
 						Integer.parseInt(plantElements[1]),
-						Integer.parseInt(plantElements[2]))));
+						Integer.parseInt(plantElements[2]),
+						0.01f, 0.1f)));
 				}
 				System.out.println("planted in 0, 0!");
 				break;
@@ -136,19 +138,19 @@ public class Main {
 		//	new Color(0.3f, 0.3f, 0.8f),
 		//	2, Plant.MAX_ITERATION)));
 
-		//plants.add(new Plant(0, 0, new Gene(15,
-		//	new LRule[] { new LRule(0.75, 'X', "Gn[[X]sXF]eG[sGX]X"),
-		//	              new LRule(0.75, 'G', "GG") },
-		//	new Color(0.0f, 0.8f, 0.2f),
-		//	new Color(0.3f, 0.3f, 0.8f),
-		//	2, Plant.MAX_ITERATION)));
+		plants.add(new Plant(0, 0, new Gene(15,
+			new LRule[] { new LRule(0.75, 'X', "Gn[[X]sXL]eG[sGXF]X"),
+			              new LRule(0.75, 'G', "GG") },
+			new Color(0.0f, 0.8f, 0.2f),
+			new Color(0.3f, 0.3f, 0.8f),
+			2, Plant.MAX_ITERATION, 0.01f, 0.1f)));
 
-		//plants.add(new Plant(1, 0, new Gene(15,
-		//	new LRule[] { new LRule(0.75, 'X', "Gs[[GF]eX]nG[wGXF]G"),
-		//	              new LRule(0.75, 'G', "GG") },
-		//	new Color(0.2f, 0.9f, 0.9f),
-		//	new Color(0.8f, 0.3f, 0.3f),
-		//	2, Plant.MAX_ITERATION)));
[…]

### 2014-03-25 src/plant_project/Gene.java
Add stemWidth and leafSize mutation

@@ -72,29 +72,52 @@ public class Gene {
 	 * mutation probability {@code Main.mutation}
 	 */
 	public Gene clone() {
+		Color   newStemColor   = stemColor;
+		Color   newFlowerColor = flowerColor;
+		float   newStemWidth   = stemWidth;
+		float   newLeafSize    = leafSize;
+		LRule[] newRules       = rules.clone();
 
-		int i = (int) (Math.random() * 100);
-		// MUTATION
-		if (i > 0 && i <= Main.mutation) {
-			LRule[] newRules = rules.clone();
-			// pick random rule to mutate
-			int mutationRuleIndex = (int) (Math.random() * (newRules.length - 1));
-			newRules[mutationRuleIndex] = mutate(newRules[mutationRuleIndex]);
-			System.out.println("old: " + rules[mutationRuleIndex]);
-			System.out.println("new: " + newRules[mutationRuleIndex]);
-			Color stemColor = new Color((float) Math.random(),
-					(float) Math.random(), (float) Math.random());
-			Color flowerColor = new Color((float) Math.random(),
-					(float) Math.random(), (float) Math.random());
-			return new Gene(angle, newRules, stemColor, flowerColor,
-					reproductionCount, maturityAge,
-					stemWidth, leafSize);
-		} else {
-			// NO MUTATION
-			return new Gene(angle, rules, stemColor, flowerColor,
-					reproductionCount, maturityAge,
-					stemWidth, leafSize);
+		// Mutation
+		if (Math.random()*100 < Main.mutation) {
+			switch ((int)(Math.random() * 5)) {
+				case 0:
+					// Mutate stem color
+					newStemColor   = new Color((float)Math.random(),
+					                           (float)Math.random(),
+					                           (float)Math.random());
+					break;
+				case 1:
+					// Mutate flower color
+					newFlowerColor = new Color((float)Math.random(),
+					                           (float)Math.random(),
+					                           (float)Math.random());
+					break;
+				case 2:
+					// Mutate stem width
+					newStemWidth = (float)(Math.random()*0.20);
+					break;
+				case 3:
+					// Mutate leaf size
+					newLeafSize  = (float)(Math.random()*0.40);
+					break;
+				case 4:
+					// Mutate rules
+					int mutationRuleIndex = (int)(Math.random()*(rules.length - 1));
+					newRules[mutationRuleIndex] = mutate(newRules[mutationRuleIndex]);
[…]

### 2014-03-25 src/plant_project/LSystem.java

@@ -1,7 +1,5 @@
 package plant_project;
 
-import java.util.Random;
-
 public class LSystem {
 	public final static char AXIOM       = 'X';
 

### 2014-03-25 src/plant_project/Main.java

@@ -19,7 +19,7 @@ public class Main {
 	/**
 	 * Mutation probability
 	 */
-	public static int mutation = 1;
+	public static int mutation = 10;
 
 	/**
 	 * Soil nutrition

### 2014-03-25 src/plant_project/Drawer.java
Start work on fitness function
Improvements:
  - Split up sun/water/nutrition fitness and make it non-linear
    E.g. having no leaves die, regardless of the rest of the fitness

Issues:
  - Leaves/flowers seem to disappear during the last round?
  - There seems to be a strong preference for large stems?
  - Fix flower/stem/leaf counts

@@ -93,7 +93,7 @@ public class Drawer extends WindowAdapter implements KeyListener,
 	}
 
 	private void drawStem(GL2 gl, Color color, float length, float radius) {
-		final int   sides  = 3;
+		final int   sides  = 8;
 
 		gl.glColor3ub((byte)color.getRed(),
 			      (byte)color.getGreen(),
@@ -203,8 +203,9 @@ public class Drawer extends WindowAdapter implements KeyListener,
 			case LSystem.GROW:
 				// drawLine
 				float lineLength = 0.1f;
-				float radius = stemWidth * stemLevel *
-					(fCount/(float)Math.sqrt(fCount))/20;
+				float radius = stemWidth;
+				//float radius = stemWidth * stemLevel *
+				//	(fCount/(float)Math.sqrt(fCount))/20;
 				this.drawStem(gl, stemColor, lineLength, radius);
 				gl.glTranslatef(0, lineLength, 0.0f);
 				break;

### 2014-03-25 src/plant_project/Gene.java

@@ -3,14 +3,22 @@ package plant_project;
 import java.awt.Color;
 
 public class Gene {
+	public static float MAX_STEM_WIDTH = 0.20f;
+	public static float MIN_STEM_WIDTH = 0.02f;
+
+	public static float MAX_LEAF_SIZE  = 0.40f;
+	public static float MIN_LEAF_SIZE  = 0.04f;
+
 	private static final char[] ORIENTATIONS = new char[] {
 		'n', 's', 'e', 'w'
 	};
+
 	private static final String[] MUTATIONS = new String[] {
 		"G", "xG", "[G]", "[xG]",
 		"F", "xF", "[F]", "[xF]",
 		"[X]", "xX",
 	};
+
 	private int angle;
 	private LRule[] rules;
 	private Color stemColor;
@@ -95,11 +103,11 @@ public class Gene {
 					break;
 				case 2:
 					// Mutate stem width
-					newStemWidth = (float)(Math.random()*0.20);
+					newStemWidth = this.randRange(MIN_STEM_WIDTH, MAX_STEM_WIDTH);
 					break;
 				case 3:
 					// Mutate leaf size
-					newLeafSize  = (float)(Math.random()*0.40);
+					newLeafSize  = this.randRange(MIN_LEAF_SIZE, MAX_LEAF_SIZE);
 					break;
 				case 4:
 					// Mutate rules
@@ -181,5 +189,9 @@ public class Gene {
 		String replace = lString.replace("x", String.valueOf(ORIENTATIONS[(int) (Math
 						.random() * (ORIENTATIONS.length - 1))]));
 		return new LRule(rule.prob, rule.find, replace);
+	}
+
+	private float randRange(float min, float max) {
+		return ((float)Math.random()) * (max-min) + min;
 	}
 }

### 2014-03-25 src/plant_project/Main.java

@@ -2,6 +2,7 @@ package plant_project;
 
 import java.awt.Color;
 import java.util.ArrayList;
+import java.util.Collections;
 import java.util.Scanner;
 
 public class Main {
@@ -9,22 +10,22 @@ public class Main {
 	/**
 	 * Water value in percent
 	 */
-	public static int water = 100;
+	public static int water = 50;
 
 	/**
 	 * Sun value in percent
 	 */
-	public static int sun = 100;
-
-	/**
-	 * Mutation probability
-	 */
-	public static int mutation = 10;
+	public static int sun = 50;
 
 	/**
 	 * Soil nutrition
 	 */
-	public static int nutrition = 10000;
+	public static int nutrition = 50;
+
+	/**
+	 * Mutation probability
+	 */
+	public static int mutation = 50;
 
 	/**
 	 * Plants in world
@@ -88,39 +89,46 @@ public class Main {
 				this.drawer.quit(); // does not return
 				break;
 			default:
-				ArrayList<Plant> newPlants = new ArrayList<Plant>(plants);
-				int n = nutrition;
-				int en;
-				for (Plant plant : plants) {
-					// plant dies
-					if (plant.getAge() >= Plant.MAX_AGE)
-						newPlants.remove(plant);
-					else {
-						// plant reproduces
-						
-						// if there's not enough nutrition, some plants die.
-						en = plant.getlString().length();
-						if (n <= en) {
-							//Delete random plants
-							//System.out.println("Deleted plant");
-							int mod = (int)(Math.random()*5);
-							for (int i=0; i < newPlants.size(); i++){
-								if (mod>2 && (i%mod)== 0){
-									Plant p = newPlants.get(i);
-									n += p.getlString().length();
-									newPlants.remove(i);
-								}
-							}
-						}
-						n -= plant.getlString().length();
-						newPlants.addAll(plant.grow());
-					}
-				}
-				plants = newPlants;
-				//System.out.println("tick");
+				// System.out.println("tick");
+				this.tick();
 				drawer.draw(plants);
 				break;
 		}
+	}
+
+	public void tick() {
+		// System.out.println("tick");
+
+		// TODO - use an interator or something instead of
+		//        doing an array copies of plants.
+		ArrayList<Plant> tmpPlants;
+
+		// TODO - make this a global
+		//        is this event the best way to do this?
+		float max_plants = 50;
+
[…]

### 2014-03-25 src/plant_project/Plant.java

@@ -4,7 +4,7 @@ import java.util.ArrayList;
 
 import plant_project.LSystem;
 
-public class Plant {
+public class Plant implements Comparable<Plant> {
 	public static final int MINUMUM_ITERATION = 4;
 	public static final int MAX_ITERATION = 7;
 	public static final int MAX_AGE = 8;
@@ -14,6 +14,7 @@ public class Plant {
 	private int age = 0;
 	private String lString = Character.toString(LSystem.AXIOM);
 	private Gene gene;
+	private float fitness;
 
 	public float getX() {
 		return x;
@@ -43,6 +44,59 @@ public class Plant {
 	}
 
 	/**
+	 * Fitness function:
+	 *
+	 * Update plant fitness based on environmental parameters.
+	 */
+	public void update(float waterPct, float sunPct, float nutritionPct) {
+		double total = 0;
+
+		// Genetic pressure from different parameters
+		//   (lots of sun == low pressure, etc)
+		double water     = 1.0 - (waterPct     / 100.0);
+		double sun       = 1.0 - (sunPct       / 100.0);
+		double nutrition = 1.0 - (nutritionPct / 100.0);
+
+		// Handy genetic properties
+		double leafRatio    = gene.getLeafSize()  / Gene.MAX_LEAF_SIZE;
+		double stemRatio    = gene.getStemWidth() / Gene.MAX_STEM_WIDTH;
+
+		double stemCount    = lString.length(); // TODO
+		double leafCount    = lString.length(); // TODO
+		double flowerCount  = lString.length(); // TODO
+
+		double storageSpace = stemRatio * stemCount;
+		double surfaceArea  = leafRatio * leafCount;
+
+		// Water
+		total -= surfaceArea    * water;      // Large surface area looses water
+		total += storageSpace/3 * water;      // Large stems can store water
+
+		// Sun
+		total += surfaceArea    * sun;        // Large leaves gather more light
+		total += stemCount      * sun;        // Taller plants have an advantage
+
+		// Nutrition
+		total -= storageSpace/2 * nutrition;  // Large stems can store nutrients
+		total -= stemCount/2    * nutrition;  // Large plants require more nutrients
+		total -= leafCount/2    * nutrition;  // ..
+		total -= flowerCount/2  * nutrition;  // ..
+
+		// Other
+		total += flowerCount;               // Flowers help with reproduction
+
[…]

### 2014-03-26 src/plant_project/Drawer.java
Add auto-play and parameter sliders

@@ -1,7 +1,11 @@
 package plant_project;
 
+import java.util.ArrayList;
+
 import java.awt.Color;
-import java.awt.Frame;
+import java.awt.BorderLayout;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
 import java.awt.event.KeyEvent;
 import java.awt.event.KeyListener;
 import java.awt.event.MouseEvent;
@@ -11,7 +15,19 @@ import java.awt.event.MouseWheelEvent;
 import java.awt.event.MouseWheelListener;
 import java.awt.event.WindowAdapter;
 import java.awt.event.WindowEvent;
-import java.util.ArrayList;
+
+import javax.swing.Timer;
+import javax.swing.JComponent;
+import javax.swing.JLabel;
+import javax.swing.JSlider;
+import javax.swing.JFrame;
+import javax.swing.JPanel;
+import javax.swing.GroupLayout;
+import javax.swing.GroupLayout.ParallelGroup;
+import javax.swing.GroupLayout.SequentialGroup;
+import javax.swing.GroupLayout.Alignment;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
 
 import javax.media.opengl.GL;
 import javax.media.opengl.GL2;
@@ -28,7 +44,8 @@ import com.google.common.base.CharMatcher;
 import com.jogamp.opengl.util.Animator;
 import com.sun.xml.internal.ws.util.StringUtils;
 
-public class Drawer extends WindowAdapter implements KeyListener,
+public class Drawer extends WindowAdapter implements
+		ActionListener, ChangeListener, KeyListener,
 		MouseListener, MouseWheelListener, MouseMotionListener,
 		GLEventListener {
 
@@ -36,6 +53,8 @@ public class Drawer extends WindowAdapter implements KeyListener,
 	public final static float EAST_BOUND  =  4.0f;
 	public final static float WEST_BOUND  = -4.0f;
 	public final static float SOUTH_BOUND = -4.0f;
+
+	public final static int   UPDATE_RATE =  10; // hz
 
 	private enum DragMode {
 		NONE,
@@ -50,10 +69,16 @@ public class Drawer extends WindowAdapter implements KeyListener,
 
 	// Interface the rest of the world
 	private Main     main;
+	private Timer    timer;
 
 	// GUI Widgets
 	private GLCanvas canvas;
[…]

### 2014-03-26 src/plant_project/Main.java

@@ -88,6 +88,8 @@ public class Main {
 			case 'q':
 				this.drawer.quit(); // does not return
 				break;
+			case ' ':
+				this.drawer.toggle();
 			default:
 				// System.out.println("tick");
 				this.tick();

### 2014-03-26 src/plant_project/Drawer.java
Randomize plant direction

@@ -299,6 +299,7 @@ public class Drawer extends WindowAdapter implements
 		gl.glPushMatrix();
 		// move to plant origin
 		gl.glTranslatef(plant.getX(), 0.0f, plant.getY());
+		gl.glRotatef(plant.getA(), 0.0f, 1.0f, 0.0f);
 
 		int fCount = CharMatcher.is(LSystem.GROW).countIn(lString);
 		float step      = 0.3f;

### 2014-03-26 src/plant_project/Gene.java

@@ -186,8 +186,9 @@ public class Gene {
 			}
 			break;
 		}
-		String replace = lString.replace("x", String.valueOf(ORIENTATIONS[(int) (Math
-						.random() * (ORIENTATIONS.length - 1))]));
+		int dir = (int)(Math.random()*ORIENTATIONS.length);
+		String replace = lString.replace("x",
+				String.valueOf(ORIENTATIONS[dir]));
 		return new LRule(rule.prob, rule.find, replace);
 	}
 

### 2014-03-26 src/plant_project/Plant.java

@@ -10,7 +10,7 @@ public class Plant implements Comparable<Plant> {
 	public static final int MAX_AGE = 8;
 	public static final float MAX_REPRODUCTION_RADIUS = 2;
 
-	private float x, y;
+	private float x, y, a;
 	private int age = 0;
 	private String lString = Character.toString(LSystem.AXIOM);
 	private Gene gene;
@@ -22,6 +22,10 @@ public class Plant implements Comparable<Plant> {
 
 	public float getY() {
 		return y;
+	}
+
+	public float getA() {
+		return a;
 	}
 
 	public int getAge() {
@@ -40,6 +44,7 @@ public class Plant implements Comparable<Plant> {
 		super();
 		this.x = x;
 		this.y = y;
+		this.a = (int)(Math.random()*360);
 		this.gene = gene;
 	}
 

### 2014-03-26 src/plant_project/Gene.java
Fix leaf mutation

@@ -16,6 +16,7 @@ public class Gene {
 	private static final String[] MUTATIONS = new String[] {
 		"G", "xG", "[G]", "[xG]",
 		"F", "xF", "[F]", "[xF]",
+		"L", "xL", "[L]", "[xL]",
 		"[X]", "xX",
 	};
 

### 2014-03-26 src/plant_project/Main.java
Shuffle plants before sorting
We assume the sort is in-place, so we want matching plants to be removed
at random.

@@ -123,6 +123,7 @@ public class Main {
 			plant.update(water, sun, nutrition);
 
 		// Sort plants and remove unfit ones
+		Collections.shuffle(plants);
 		Collections.sort(plants);
 		while (plants.size() > max_plants)
 			plants.remove(0);

### 2014-03-26 src/plant_project/Plant.java
Tweak fitness function
This still isn't the greatest, it's very linear, but maybe it's better
than it was?

@@ -2,7 +2,7 @@ package plant_project;
 
 import java.util.ArrayList;
 
-import plant_project.LSystem;
+import com.google.common.base.CharMatcher;
 
 public class Plant implements Comparable<Plant> {
 	public static final int MINUMUM_ITERATION = 4;
@@ -54,8 +54,6 @@ public class Plant implements Comparable<Plant> {
 	 * Update plant fitness based on environmental parameters.
 	 */
 	public void update(float waterPct, float sunPct, float nutritionPct) {
-		double total = 0;
-
 		// Genetic pressure from different parameters
 		//   (lots of sun == low pressure, etc)
 		double water     = 1.0 - (waterPct     / 100.0);
@@ -66,31 +64,39 @@ public class Plant implements Comparable<Plant> {
 		double leafRatio    = gene.getLeafSize()  / Gene.MAX_LEAF_SIZE;
 		double stemRatio    = gene.getStemWidth() / Gene.MAX_STEM_WIDTH;
 
-		double stemCount    = lString.length(); // TODO
-		double leafCount    = lString.length(); // TODO
-		double flowerCount  = lString.length(); // TODO
+		double stemCount    = CharMatcher.is(LSystem.GROW).countIn(lString);
+		double leafCount    = CharMatcher.is(LSystem.LEAF).countIn(lString);
+		double flowerCount  = CharMatcher.is(LSystem.FLOWER).countIn(lString);
 
 		double storageSpace = stemRatio * stemCount;
 		double surfaceArea  = leafRatio * leafCount;
+		double biomass      = storageSpace*2 + surfaceArea + flowerCount*2;
+
+		// Fitness
+		double wfit = 0;
+		double sfit = 0;
+		double nfit = 0;
 
 		// Water
-		total -= surfaceArea    * water;      // Large surface area looses water
-		total += storageSpace/3 * water;      // Large stems can store water
+		wfit  -= surfaceArea    * water;      // Large surface area looses water
+		wfit  += stemRatio      * water;      // Large stems can store water
 
 		// Sun
-		total += surfaceArea    * sun;        // Large leaves gather more light
-		total += stemCount      * sun;        // Taller plants have an advantage
+		sfit  += surfaceArea*2  * sun;        // Large leaves gather more light
+		sfit  += stemCount      * sun;        // Taller plants have an advantage
 
 		// Nutrition
-		total -= storageSpace/2 * nutrition;  // Large stems can store nutrients
[…]

### 2014-03-27 src/plant_project/Drawer.java
Add screenshot saving

@@ -3,6 +3,7 @@ package plant_project;
 import java.util.ArrayList;
 
 import java.awt.Color;
+import java.awt.Component;
 import java.awt.BorderLayout;
 import java.awt.event.ActionEvent;
 import java.awt.event.ActionListener;
@@ -15,6 +16,7 @@ import java.awt.event.MouseWheelEvent;
 import java.awt.event.MouseWheelListener;
 import java.awt.event.WindowAdapter;
 import java.awt.event.WindowEvent;
+import java.awt.image.BufferedImage;
 
 import javax.swing.Timer;
 import javax.swing.JComponent;
@@ -34,7 +36,7 @@ import javax.media.opengl.GL2;
 import javax.media.opengl.GL2ES1;
 import javax.media.opengl.GLAutoDrawable;
 import javax.media.opengl.GLEventListener;
-import javax.media.opengl.awt.GLCanvas;
+import javax.media.opengl.awt.GLJPanel;
 import javax.media.opengl.fixedfunc.GLLightingFunc;
 import javax.media.opengl.fixedfunc.GLMatrixFunc;
 import javax.media.opengl.glu.GLU;
@@ -72,7 +74,7 @@ public class Drawer extends WindowAdapter implements
 	private Timer    timer;
 
 	// GUI Widgets
-	private GLCanvas canvas;
+	private GLJPanel canvas;
 	private JPanel   tools;
 	private JSlider  water;
 	private JSlider  sun;
@@ -90,7 +92,7 @@ public class Drawer extends WindowAdapter implements
 	public Drawer(Main main) {
 		this.main      = main;
 		this.timer     = new Timer(1000/UPDATE_RATE, this);
-		this.canvas    = new GLCanvas();
+		this.canvas    = new GLJPanel();
 		this.tools     = new JPanel();
 		this.water     = new JSlider(0, 100);
 		this.sun       = new JSlider(0, 100);
@@ -433,6 +435,17 @@ public class Drawer extends WindowAdapter implements
 		System.out.println(running ? "stopped" : "started");
 	}
 
+	public BufferedImage toImage(boolean full) {
+		this.canvas.setupPrint(1, 1, 0, -1, -1);
+		Component comp = full ? this.frame : this.canvas;
+		BufferedImage image = new BufferedImage(
+			comp.getWidth(),
+			comp.getHeight(),
+			BufferedImage.TYPE_INT_RGB);
[…]

### 2014-03-27 src/plant_project/Main.java

@@ -1,9 +1,16 @@
 package plant_project;
 
-import java.awt.Color;
+import java.io.File;
+
 import java.util.ArrayList;
 import java.util.Collections;
+import java.util.Date;
 import java.util.Scanner;
+
+import java.awt.Color;
+import java.awt.image.BufferedImage;
+
+import javax.imageio.ImageIO;
 
 public class Main {
 
@@ -88,6 +95,9 @@ public class Main {
 			case 'q':
 				this.drawer.quit(); // does not return
 				break;
+			case 'c':
+				this.screenshot(); // take screen capture
+				break;
 			case ' ':
 				this.drawer.toggle();
 			default:
@@ -98,6 +108,20 @@ public class Main {
 		}
 	}
 
+	public void screenshot() {
+		final Date time = new Date();
+		final String path = String.format("img/screenshot-%tY%tm%td_%tH%tM%tS_%tL.png",
+				time, time, time, time, time, time, time);
+		System.out.println("Saving - " + path);
+		try {
+			File file = new File(path);
+			file.mkdirs();
+			BufferedImage img = this.drawer.toImage(true);
+			ImageIO.write(img, "png", file);
+		} catch(Exception e) {
+			e.printStackTrace();
+		}
+	}
 	public void tick() {
 		// System.out.println("tick");
 

### 2014-03-27 src/plant_project/Drawer.java
Add video recording
This seems to only work for small video sizes

@@ -1,6 +1,8 @@
 package plant_project;
 
 import java.util.ArrayList;
+import java.io.File;
+import javax.imageio.ImageIO;
 
 import java.awt.Color;
 import java.awt.Component;
@@ -72,6 +74,7 @@ public class Drawer extends WindowAdapter implements
 	// Interface the rest of the world
 	private Main     main;
 	private Timer    timer;
+	private Video    video;
 
 	// GUI Widgets
 	private GLJPanel canvas;
@@ -443,11 +446,45 @@ public class Drawer extends WindowAdapter implements
 			comp.getHeight(),
 			BufferedImage.TYPE_INT_RGB);
 		comp.printAll(image.getGraphics());
+		this.canvas.releasePrint();
 		return image;
+	}
+
+	public void capture(String path) {
+		try {
+			System.out.println("Saving - " + path);
+			File file = new File(path);
+			file.getParentFile().mkdirs();
+			BufferedImage img = this.toImage(true);
+			ImageIO.write(img, "png", file);
+		} catch(Exception e) {
+			e.printStackTrace();
+		}
+	}
+
+	public void record(String path) {
+		if (this.video == null) {
+			System.out.println("Record - " + path);
+			File file = new File(path);
+			file.getParentFile().mkdirs();
+			this.video = new Video(file, UPDATE_RATE);
+		} else {
+			System.out.println("Record - done");
+			Video vid = this.video;
+			this.video = null;
+			vid.finish();
+		}
 	}
 
 	public void draw(ArrayList<Plant> plants) {
 		this.plants = plants;
+
+		// record video
+		if (this.video != null) {
+			System.out.println("Record - frame");
+			BufferedImage img = this.toImage(true);
+			this.video.encodeImage(img);
+		}
 	}
 
 	public void pan(double fwd, double side, double up) {

### 2014-03-27 src/plant_project/Main.java

@@ -1,16 +1,10 @@
 package plant_project;
 
-import java.io.File;
-
+import java.awt.Color;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.Date;
 import java.util.Scanner;
-
-import java.awt.Color;
-import java.awt.image.BufferedImage;
-
-import javax.imageio.ImageIO;
 
 public class Main {
 
@@ -49,6 +43,8 @@ public class Main {
 	 */
 	public void keyPress(char c, String in)
 	{
+		Date   time;
+		String path;
 		switch (c) {
 			case 'p':
 				// plant p=(probability:find:replace; ...), reproducecount, maturityage
@@ -96,7 +92,20 @@ public class Main {
 				this.drawer.quit(); // does not return
 				break;
 			case 'c':
-				this.screenshot(); // take screen capture
+				// take screen capture
+				time = new Date();
+				path = String.format(
+					"img/screenshot-%tY%tm%td_%tH%tM%tS_%tL.png",
+					time, time, time, time, time, time, time);
+				this.drawer.capture(path);
+				break;
+			case 'r':
+				// toggle video recording
+				time = new Date();
+				path = String.format(
+					"img/video-%tY%tm%td_%tH%tM%tS.mp4",
+					time, time, time, time, time, time);
+				this.drawer.record(path);
 				break;
 			case ' ':
 				this.drawer.toggle();
@@ -108,20 +117,6 @@ public class Main {
 		}
 	}
 
-	public void screenshot() {
-		final Date time = new Date();
-		final String path = String.format("img/screenshot-%tY%tm%td_%tH%tM%tS_%tL.png",
-				time, time, time, time, time, time, time);
-		System.out.println("Saving - " + path);
-		try {
-			File file = new File(path);
-			file.mkdirs();
-			BufferedImage img = this.drawer.toImage(true);
-			ImageIO.write(img, "png", file);
-		} catch(Exception e) {
-			e.printStackTrace();
-		}
-	}
 	public void tick() {
 		// System.out.println("tick");
 

### 2014-03-27 src/plant_project/Video.java

@@ -1,0 +1,106 @@
+// http://stackoverflow.com/questions/11006910/encode-a-video-into-h264-using-bufferedimages
+
+package plant_project;
+
+import java.util.ArrayList;
+import java.nio.ByteBuffer;
+import java.io.File;
+import java.io.IOException;
+import java.awt.image.BufferedImage;
+
+import org.jcodec.common.NIOUtils;
+import org.jcodec.common.SeekableByteChannel;
+import org.jcodec.common.model.ColorSpace;
+import org.jcodec.common.model.Picture;
+import org.jcodec.codecs.h264.H264Encoder;
+import org.jcodec.codecs.h264.H264Utils;
+import org.jcodec.containers.mp4.Brand;
+import org.jcodec.containers.mp4.MP4Packet;
+import org.jcodec.containers.mp4.TrackType;
+import org.jcodec.containers.mp4.muxer.FramesMP4MuxerTrack;
+import org.jcodec.containers.mp4.muxer.MP4Muxer;
+import org.jcodec.scale.AWTUtil;
+import org.jcodec.scale.RgbToYuv420;
+
+public class Video {
+	private SeekableByteChannel ch;
+	private Picture toEncode;
+	private RgbToYuv420 transform;
+	private H264Encoder encoder;
+	private ArrayList<ByteBuffer> spsList;
+	private ArrayList<ByteBuffer> ppsList;
+	private FramesMP4MuxerTrack outTrack;
+	private ByteBuffer _out;
+	private int frameNo;
+	private MP4Muxer muxer;
+	private int fps;
+
+	public Video(File out, int fps) {
+		try {
+			this.ch = NIOUtils.writableFileChannel(out);
+			this.fps = fps;
+
+			// Transform to convert between RGB and YUV
+			transform = new RgbToYuv420(0, 0);
+
+			// Muxer that will store the encoded frames
+			muxer = new MP4Muxer(ch, Brand.MP4);
+
+			// Add video track to muxer
+			outTrack = muxer.addTrackForCompressed(TrackType.VIDEO, fps);
+
+			// Allocate a buffer big enough to hold output frames
+			_out = ByteBuffer.allocate(1920 * 1080 * 6);
+
+			// Create an instance of encoder
+			encoder = new H264Encoder();
+
+			// Encoder extra data ( SPS, PPS ) to be stored in a special place of
+			// MP4
+			spsList = new ArrayList<ByteBuffer>();
[…]

### 2014-03-27 src/plant_project/Main.java
Add simple example plant

@@ -168,6 +168,13 @@ public class Main {
 		//	new Color(0.3f, 0.3f, 0.8f),
 		//	2, Plant.MAX_ITERATION)));
 
+		//plants.add(new Plant(0, 0, new Gene(15,
+		//	new LRule[] { new LRule(0.75, 'X', "GGGG[nX][sX][eX][wX][X]"),
+		//	              new LRule(0.75, 'G', "GG") },
+		//	new Color(0.0f, 0.8f, 0.2f),
+		//	new Color(0.3f, 0.3f, 0.8f),
+		//	2, Plant.MAX_ITERATION, 0.01f, 0.1f)));
+
 		plants.add(new Plant(0, 0, new Gene(15,
 			new LRule[] { new LRule(0.75, 'X', "Gn[[X]sXL]eG[sGXF]X"),
 			              new LRule(0.75, 'G', "GG") },

### 2014-03-27 doc/paper.tex
Work on the report

@@
 \usepackage{enumitem}
+\usepackage[utf8]{inputenc}
 
@@
   \qquad
-  Andy Spencer\thanks{e-mail:e-mail:andy753421@ucla.edu}
+  Andy Spencer\thanks{e-mail:andy753421@ucla.edu}
   \qquad
@@
 %%%%%%%%%%%%
+
+% Proposal
+% --------
+% Team:
+%
+%   Andy Philipp <philipp.andy@gmail.com>
+%   Andy Spencer <andy753421@ucla.edu>
+%   Haakon Garseg Moerk <haakongmork@gmail.com>
+%
+% Title:
+%
+%   Plant Modeling Using Environmental Parameters
+%
+% Summary:
+%
+%   The project will focus on modeling natural looking plants using
+%   simulated environment parameters to control growth and evolution of
+%   different plant species. Genetic algorithms or other evolutionary
+%   techniques will be used to evolve plant suitable for growth in various
+%   environments.
+%
+% Potential environmental conditions:
+%
+%   - Availability of water
+%   - Amount of sunlight
+%   - Soil conditions
+%   - Growth elevation
+%
+% Potential plant species parameters:
+%
+%   - Rate of water loss
+%   - Food and water storage capability
+%   - Total mass/volume
+%   - Crown height
+%   - Leaf area
+%   - Root structures
+
+% Project submission
+% ------------------
+% We need to make web page containing:
+%
+%   1. The project abstract
+%   2. A link to the paper
+%   3. Links to any demos or videos
+%
+% The paper should be in conference format:
+%
+%   1. We can use two-column style with, e.g.:
+%        a) references,  b) technical description,
+%        c) results,     d) conclusions
+%   2. It should be medium length, 5 pages is fine.
+
+% Outline
+% -------
+% 1. Introduction
+% 2. Architecture
+% 3. Plant System
+% 4. Genetics
+% 5. Results
+% 6. Conclusions
 
@@
 
-\section{Architecture and Models}
+\section{Software Architecture}
+
+The software was implemented using Java and OpenGL for drawing. The top-level
+architecture can be divided into two main categories: plant simulation using
+L-Systems and the graphical user interface.
+
+\subsection{Plants}
+
+The L-Systems are implemented in the LSystem and LRule classes, which are
+responsible for storing the production rules associated with each plant and
+iterating the plants phenotype as the plant grows throughout the simulation. The
+Gene and Plant classes store representations of each living plant. The Gene
+class represents the plant genotype as a combination of the L-System production
+rules along with additional information such as stem and leaf size, stem and
+flower colors, the number of seeds produced, and the age at which the plant
+matures. Each individual plants phenotype is represented using the Plant class
+and contains it's X/Y position and orientation inside the world, along with the
+current age and a L-System string (L-String) representing the plants current
+shape. The Plant class is also responsible for calculating each individual
+plants fitness values at the end of each simulation step.
+
+\subsection{User Interface}
+
+The graphical interface is comprised of the Main class, a Drawer class, and a
+Video utility class for recording screen captures of the simulation. The Main
+class receives user input to control the plant simulation. It also contains the
+environmental parameters and invokes the plant growth and fitness functions. The
+Drawer class receives a list of plants from the Main class and displays them to
+the user. The graphical interfaces also provides convenient controls for each of
+the environmental parameters so that they can be tuned while the simulation is
+executing.
 
@@
 
-\section{Simulation}
+\section{Plant Simulation}
 
-The simulation of the plant environment is governed by ticks. For each tick that
-goes, all plant objects becomes older. We wanted to have a realistic approach to
-have plant grow in the real world, and therefore the first 4 ticks, the plant is
-growing as a seed under ground. On the fifth tick it starts to grow above
-ground. When the seventh tick is reached, the plant becomes mature and grows
-flowers and leaves, if the gene is capable of it. The plant object then spreads
-it seeds within a certain radius. When the season is over, the plant dies, but
-it’s children grows up with either the same gene, or and mutation of the gene.
+Each plant in our environment contains a string of symbol (which we call an
+"L-String") representing the plants current shape and size. Each symbol can
+represent a stem segment, a flower, a leaf, etc. The plants genotype contains a
+set of production rules that govern the growth and development of the plant.
+When the plant grows, the production rules are used to replace symbols in the
+plants current L-String with sequences of tokens to produces a larger and more
+complex plant.
 
-\section{Genetics}
+\subsection{L-System Genotypes}
 
-In plant biology it’s defined two types of reproduction. Asexual reproduction
-and sexual reproduction. Asexual reproduction can either be a clone or a
-mutation of its parent. Sexual reproduction creates a new offspring by combining
-the genetics from its parents.\cite{plantrepo}
-
-\section{Conclusion}
-
-Gene:
+We implemented a modified variant of L-Systems where a replacement probability
+is included in addition to the matching and replacement parts of each production
+rule. The possible values for each symbol in our rules are as follows:
 
 \begin{description}[leftmargin=!,labelindent=0.2in,labelwidth=0.1in]
-  \item[F]   grow stem
-  \item[L]   grow flower
+  \item[G]   grow a stem segment
+  \item[F]   grow a flower at the current position
+  \item[L]   grow a leaf at the current position
   \item[X]   does not correspond to any drawing action and is used to control
@@
              The angle is dependent on the genes angle.
-  \item[{[}] push matrix
-  \item[{]}] pop matrix
+  \item[{[}] push matrix, the current position and angle of growth are saved.
+  \item[{]}] pop matrix, the position and angle from he corresponding {]} are
+             restored.
 \end{description}
 
-Example of a plant at first generation:
+Using the push and pop symbols, {[} and {]}, the plant can grow branches in
+different directions. For example, a north -- south Y shape can be grown using
+the string: \[ G [ s G ] [ n G ] \]
+
+An example of a plant at first generation:
 
@@
   \item[Angle] 15 degrees
-  \item[Axiom] X
-  \item[Rules] \{ ``Fs[[F]eX]nF[wFX]F'' , ``FF'' \}
+  \item[Axiom] $ X $
+  \item[Rules] $ X \rightarrow Gs[[G]eX]nG[wGX]G \\
+                 G \rightarrow GG $
 \end{description}
 
-Notice that this plant will not start with flowers (`L').
+Notice that this plant will not start with flowers ($F$) or leaves ($L$).
 
-Generally when the plant grows, it takes a random character from the LString
-(but not `[' or `]') and substitutes this character with one random rule from
-the plant gene. At the seventh tick of the season, the plant is mature and will
-clone itself. There's two ways that this can happen. Either the new plant will
-grow up with the same gene, or the gene will mutate and the new plant will be
-given the mutated gene. The mutation happens in the set of the gene rules. A
-random rule is picked out to be mutated. Again there is two possible outcomes.
-Either a character in the rule will be deleted, or a random string from a
-predefined mutation array will be added at a random position in the rule. We
-want to simulate subtle changes in the gene as the mutation isn't suppose to be
-extreme in any ways. Therefore the mutation array can be in the following way \{
-``F'', ``L'', ``xF'', ``xL'', ``xX'', ``[X]'', ``[F]'', ``[L]'', ``[xF]'',
-``[xL]'' \}. `x' is substituted with a random direction to grow, i.e. `n', `e',
-`s' or `w'.
+\subsection{Simulation}
 
-\section{Environmental parameters}
+The simulation of the plant environment is governed by ticks. During each tick,
+each of the currently living plant objects becomes older. We wanted to have a
+similar approach to have plants grow in the real world. Consequently, during the
+first 4 ticks the plant grows as a seed under ground. On the fifth tick it
+starts growing above ground. When the seventh tick is reached, the plant becomes
+mature and grows flowers and leaves, if the plants genotype is capable of it.
+The plant object then spreads it seeds within a certain radius. When the season
+is over, the plant dies, but it's children grows up with either the same gene,
+or a mutation of the gene.
+
+\section{Genetic Algorithms}
+
+In plant biology, reproduction is divided into two categories: asexual
+reproduction and sexual reproduction. In our project we chose to use asexual
+reproduction to avoid implementing genetic crossover and recombination. Each of
+our plants is initially a clone of it's parent, however there is a probability
+that mutations will be introduced during the cloning process.\cite{plantrepo}
+
+In our simulation, the seasons are divided into ticks, at the seventh tick of
+the season, the plants are considered mature and will produce seeds. After the
+plants have reproduced and the season has ended all the existing plants are
+removed and the new child plants are given a chance to grow.
+
+\subsection{Mutation}
+
+When each plant produces seeds, the seed may mutation in several different
+ways. Mutation is controlled based on a mutation probability global to all the
+plants, if the seed is chosen to be mutated it will change randomly in one of
+the following ways:
+
+\begin{description}
+  \item[Stem color]   \hfill \\
+    The plants stem color is changed randomly, the stem color color is also used
+    for the plants leaves.
+  \item[Flower color] \hfill \\
+    The plants flower color is changed randomly. Both the stem color and flower
+    color are only used for display and can be used to help differentiate plants
+    of different species.
+  \item[Stem width]   \hfill \\
+    The diameter of the plants stem is changed. Upper and lower limits are
+    imposed and a width is randomly chosen from within the range.
+  \item[Leaf size]    \hfill \\
+    The length of each leaf is changed. Upper and lower limits are used. The
+    leaf shape is fixed for drawing purposes.
+  \item[Growth rules] \hfill \\
+    The L-System productions used to control the plants growth are changed. The
+    plant may become either more or less complex.
+\end{description}
+
+\subsubsection{Growth rule mutation}
+
+We want to simulate subtle changes in the gene as the mutation isn't suppose to
+be extreme in any ways. Gene mutation occurs by choosing one of the production
+rules and then mutation the replacement side of the rule. A symbol is randomly
+chosen from the replacement to be mutated.
+
+The mutation can either be that the symbols is simply deleted, or that is then
+replaced with a sequences from a predefined set of possible mutations.
+
+The replacement sequences we used are logically divided into several logical
+groups, however these divisions are just for convenience and are not used by the
+mutation algorithm.
+
+\begin{description}[leftmargin=!,labelindent=0.2in,labelwidth=0.5in]
+  \item[Stems]   $G,~~xG,~~[G],~~[xG]$
+  \item[Flowers] $F,~~xF,~~[F],~~[xF]$
+  \item[Leaves]  $L,~~xL,~~[L],~~[xL]$
+  \item[Other]   $[X],~~xX$
+\end{description}
+
+Note that in in each of the replacements, `x' is substituted with a random
+direction to grow, i.e. `n', `e', `s' or `w'.
+
+\subsection{Fitness}
+
+% // Water
+% wfit  -= surfaceArea   * water;      // Large surface area looses water
+% wfit  += stemRatio     * water;      // Large stems can store water
+% 
+% // Sun
+% sfit  += surfaceArea*2 * sun;        // Large leaves gather more light
+% sfit  += stemCount     * sun;        // Taller plants have an advantage
+% 
+% // Nutrition
+% nfit  -= stemRatio     * nutrition;  // Large plants require more nutrients
+% nfit  -= surfaceArea   * nutrition;  // Large plants require more nutrients
+% nfit  -= stemCount     * nutrition;  // Large plants require more nutrients
+
+\subsubsection{Environmental parameters}
 
@@
 in the ground.
-
-\section{Graphics}
 

### 2014-03-28 src/plant_project/Drawer.java
Simulation and fitness tweaks

@@ -300,6 +300,7 @@ public class Drawer extends WindowAdapter implements
 		Color flowerColor = plant.getGene().getFlowerColor();
 		float stemWidth   = plant.getGene().getStemWidth();
 		float leafSize    = plant.getGene().getLeafSize();
+		float maturityAge = plant.getGene().getMaturityAge();
 
 		gl.glPushMatrix();
 		// move to plant origin
@@ -328,8 +329,11 @@ public class Drawer extends WindowAdapter implements
 				this.drawLeaf(gl, stemColor, leafSize);
 				break;
 			case LSystem.FLOWER:
-				if (plant.getAge() >= Plant.MAX_ITERATION)
+				if (plant.getAge() >= maturityAge-1) {
+					gl.glTranslatef( stemWidth, 0f, 0f);
 					this.drawFlower(gl, flowerColor);
+					gl.glTranslatef(-stemWidth, 0f, 0f);
+				}
 				break;
 			case LSystem.PUSH:
 				// push position and angle

### 2014-03-28 src/plant_project/Main.java

@@ -9,6 +9,11 @@ import java.util.Scanner;
 public class Main {
 
 	/**
+	 * Maximum number of plants
+	 */
+	public static final int MAX_PLANTS = 30;
+
+	/**
 	 * Water value in percent
 	 */
 	public static int water = 50;
@@ -118,15 +123,30 @@ public class Main {
 	}
 
 	public void tick() {
-		// System.out.println("tick");
-
 		// TODO - use an interator or something instead of
 		//        doing an array copies of plants.
 		ArrayList<Plant> tmpPlants;
+		int doomed = 0;
 
-		// TODO - make this a global
-		//        is this event the best way to do this?
-		float max_plants = 50;
+		// Only run fitness for gorwing plants
+		if (plants.get(0).getAge() >= Plant.MIN_ITERATION) {
+			doomed = (plants.size() - MAX_PLANTS) / 2;
+
+			// Run fitness function
+			for (Plant plant : plants)
+				plant.update(water, sun, nutrition);
+
+			// Sort plants and remove unfit ones
+			Collections.shuffle(plants);
+			Collections.sort(plants);
+			for (int i = 0; i < doomed; i++)
+				plants.remove(0);
+		}
+
+		// Add offspring
+		tmpPlants = new ArrayList<Plant>(plants);
+		for (Plant plant : tmpPlants)
+			plants.addAll(plant.grow());
 
 		// Kill off old plants
 		tmpPlants = new ArrayList<Plant>(plants);
@@ -134,23 +154,8 @@ public class Main {
 			if (plant.getAge() >= Plant.MAX_AGE)
 				plants.remove(plant);
 
-		// Run fitness function
-		//float num = (float)plants.size();
-		//for (Plant plant : plants)
-		//	plant.update(water/num, sun/num, nutrition/num);
-		for (Plant plant : plants)
-			plant.update(water, sun, nutrition);
-
-		// Sort plants and remove unfit ones
-		Collections.shuffle(plants);
-		Collections.sort(plants);
-		while (plants.size() > max_plants)
-			plants.remove(0);
-
-		// Add offspring
-		tmpPlants = new ArrayList<Plant>(plants);
-		for (Plant plant : tmpPlants)
-			plants.addAll(plant.grow());
+		// Debug
+		System.out.println("tick - removed " + doomed);
 	}
 
 	/**

### 2014-03-28 src/plant_project/Plant.java

@@ -5,9 +5,9 @@ import java.util.ArrayList;
 import com.google.common.base.CharMatcher;
 
 public class Plant implements Comparable<Plant> {
-	public static final int MINUMUM_ITERATION = 4;
-	public static final int MAX_ITERATION = 7;
-	public static final int MAX_AGE = 8;
+	public static final int   MIN_ITERATION = 4;
+	public static final int   MAX_ITERATION = 8;
+	public static final int   MAX_AGE       = 8;
 	public static final float MAX_REPRODUCTION_RADIUS = 2;
 
 	private float x, y, a;
@@ -76,24 +76,29 @@ public class Plant implements Comparable<Plant> {
 		double wfit = 0;
 		double sfit = 0;
 		double nfit = 0;
+		double efit = 0;
 
 		// Water
 		wfit  -= surfaceArea    * water;      // Large surface area looses water
-		wfit  += stemRatio      * water;      // Large stems can store water
+		wfit  += storageSpace   * water;      // Large stems can store water
 
 		// Sun
 		sfit  += surfaceArea*2  * sun;        // Large leaves gather more light
 		sfit  += stemCount      * sun;        // Taller plants have an advantage
 
 		// Nutrition
-		nfit  -= stemRatio      * nutrition;  // Large plants require more nutrients
-		nfit  -= surfaceArea    * nutrition;  // Large plants require more nutrients
-		nfit  -= stemCount      * nutrition;  // Large plants require more nutrients
+		nfit  -= surfaceArea  * nutrition;  // Large plants require more nutrients
+		nfit  -= storageSpace * nutrition;  // ..
+		nfit  -= stemCount    * nutrition;  // ..
 
 		// Other
 		//total += flowerCount;                 // Flowers help with reproduction
 
-		this.fitness = (float)(wfit+sfit+nfit);
+		// Sanity checkds
+		if (stemCount == 0)
+			efit = -100;
+
+		this.fitness = (float)(wfit+sfit+nfit+efit);
 
 		System.out.format("%7.2f + %7.2f + %7.2f = %7.2f\n",
 				wfit, sfit, nfit, this.fitness);
@@ -124,7 +129,7 @@ public class Plant implements Comparable<Plant> {
 
 		ArrayList<Plant> children = new ArrayList<Plant>();
 		++age;
-		if (MINUMUM_ITERATION <= age && age <= MAX_ITERATION) {
[…]

### 2014-03-28 doc/paper.tex
Finish initial paper draft

@@
 \usepackage{enumitem}
+\usepackage{amsmath}
 \usepackage[utf8]{inputenc}
@@
 shape. The Plant class is also responsible for calculating each individual
-plants fitness values at the end of each simulation step.
+plants fitness values at the end of each simulation tick.
 
@@
 
+At the start of each simulation tick, each existing plant updates a fitness
+value based on the current environmental parameters and the plants current
+growth state. The list of plants is then sorted based on fitness and unfit
+plants are removed if the environment is over crowded.
+
 \subsubsection{Environmental parameters}
 
-There are three basic environmental parameters which has an effect on the
-plants. The amount of sun (0-100\%), water (0-100\%) and the nutrition available
-in the ground.
+There are three basic environmental parameters which have an effect on the
+plants.
+
+\begin{description}[leftmargin=!,labelindent=0.2in,labelwidth=0.5in]
+  \item[sun]       The amount of sunlight
+  \item[water]     The availability of water
+  \item[nutrition] The nutritional content of the soil
+\end{description}
+
+Each of these parameters initially ranges from 0 to 100 and can be changed by
+the user interacting with the software. In addition, a mutation parameter
+controls the rate of change in the plant genomes.
+
+For the purpose of fitness calculations, a ``genetic pressure'' version of each
+of these parameters is calculated and used in the fitness function.
+\[pressure = 1 - (parameter / 100)\]
+
+\subsubsection{Plant parameters}
+
+When calculating each plants fitness several plant parameters are used:
+
+\begin{description}[leftmargin=!,labelindent=0.2in,labelwidth=0.1in]
+  \item[Stem ratio]    \hfill \\
+    Normalized stem width, $0 .. 1$
+  \item[Leaf ratio]    \hfill \\
+    Normalized leaf size, $0 .. 1$
+  \item[Stem count]    \hfill \\
+    The number of stem segments the plant has grown thus far.
+  \item[Leaf count]    \hfill \\
+    The number of leaves grown thus far.
+  \item[Storage space] \hfill \\
+    The total storage space available to the plant. We do not calculate an
+    actual volume but instead just use a combination of the stem width from the
+    plants genes and the number of stem segments.
+    \[ storageSpace = stemRatio \times stemCount \]
+  \item[Surface area]  \hfill \\
+    The total surface area available in the leaves for photosynthesis. Like the
+    storage space, this is a combination of the leaf size and the number of
+    leaves.
+    \[ surfaceArea = leafRatio \times leafCount \]
+\end{description}
+
+\subsubsection{Fitness calculation}
+
+The plant fitness function is divided by water, sun, an nutrient fitness. Each
+of these is calculated separately and the total fitness is the sum. We
+implemented the fitness equations such that if all the environmental parameters
+are equal, then all the fitness values will also be equal and the plants will
+evolve randomly. In a more comprehensive system, some plants could be considered
+more or less fit in any environment including an environment with equal
+environmental parameters.
+
+The following guidelines were used by our fitness function:
+
+\paragraph{Water fitness}
+\begin{itemize}
+  \item Large stems (high storage space) can conserve water and result in higher
+    fitness in dry environments.
+  \item A large surface area results in larger water loss and a lower
+    fitness in dry environments.
+\end{itemize}
+\vspace{-0.5em}
+\begin{eqnarray*}
+waterFitness &=& (storageSpace \times waterPressure) \\
+             &-& (surfaceArea \times waterPressure)
+\end{eqnarray*}
+
+\paragraph{Sunlight fitness}
+\begin{itemize}
+  \item A large surface area collects more light and results in a higher fitness
+    in shady environments.
+  \item More stem segments will result in a taller plant which can college more
+    light. We do not calculate actual heights, but assume that stem segments
+    improve light gathering.
+\end{itemize}
+\vspace{-0.5em}
+\begin{eqnarray*}
+sunFitness &=& (2 \times surfaceArea \times sunPressure) \\
+           &+& (stemCount \times sunPressure)
+\end{eqnarray*}
+
+\paragraph{Nutrition fitness}
+\begin{itemize}
+  \item Larger stems, more stem segments, and a larger surface all result in
+    higher nutrient requirements and thus lower the fitness in low quality soil.
+\end{itemize}
+\vspace{-0.5em}
+\begin{eqnarray*}
+nutrientFitness &=&-(surfaceArea \times nutrientPressure) \\
+                &+&-(storageSpace \times nutrientPressure) \\
+                &+&-(storageSpace \times nutrientPressure)
+\end{eqnarray*}
 
 \section{Conclusion}
+
+During our simulations we were able to adjust the environmental parameters in
+order to control the growth and evolution of different plant species. In many
+cases the evolved plants share some resemblance with natural plants found in
+similar environments. Some example of this include stout leafless plants
+resembling cacti which evolve in dry environments, large leafy plants resembling
+trees that evolve in shady environments, and small shrubby or grass like plants
+that evolve in poor soil conditions.
+
+In future programs we would recommend improving the genetic fitness functions by
+taking into consideration environmental information related to the neighboring
+plants in the environment. This would use more advanced lighting and collision
+detection to calculate sunlight based on shadows in the environment.
+
+In addition, more realistic fitness levels could be used that are not linear
+with regards to the amount of pressure from different environmental parameters.
+In our system, a leafy plant may always win over other plants in a shady
+simulation environment even though such as large pant may not be support in a
+physical environment to do water or soil limitations.
+
+Some difficulty was encountered evolving plants to appear realistic, and many
+limitations are inherit which are difficult to simulation in a computer graphics
+setting. For example, two physical plants which share a remarkable resemblances
+my be suited to different environments due to small evolutionary changes. Waxy
+leaves, for example, could result in a leafy plant well suited to a dry
+environment.
+
+Overall, we have shown that artificial plants based on L-Systems can be evolved
+in realistic ways based on abstract environmental parameters. Using L-Strings to
+represent plants results in both realistic looking shapes and provides the
+foundation for genetic algorithm for controlling the development of plant
+species.
 

### 2014-03-28 src/plant_project/Drawer.java
Improve video record support

@@ -94,6 +94,7 @@ public class Drawer extends WindowAdapter implements
 
 	public Drawer(Main main) {
 		this.main      = main;
+		this.plants    = main.plants;
 		this.timer     = new Timer(1000/UPDATE_RATE, this);
 		this.canvas    = new GLJPanel();
 		this.tools     = new JPanel();
@@ -480,12 +481,9 @@ public class Drawer extends WindowAdapter implements
 		}
 	}
 
-	public void draw(ArrayList<Plant> plants) {
-		this.plants = plants;
-
-		// record video
+	public void addFrame(String id) {
 		if (this.video != null) {
-			System.out.println("Record - frame");
+			System.out.println("Record - frame - " + id);
 			BufferedImage img = this.toImage(true);
 			this.video.encodeImage(img);
 		}
@@ -510,6 +508,9 @@ public class Drawer extends WindowAdapter implements
 		while (this.location[1] < -180) this.location[1] += 360;
 		while (this.location[1] >  180) this.location[1] -= 360;
 		this.location[2] = Math.abs(this.location[2]);
+
+		/* Record frame */
+		this.addFrame("pan");
 	}
 
 	public void rotate(double x, double y, double z) {
@@ -527,12 +528,15 @@ public class Drawer extends WindowAdapter implements
 		while (this.rotation[1] >  180) this.rotation[1] -= 360;
 		while (this.rotation[2] < -180) this.rotation[2] += 360;
 		while (this.rotation[2] >  180) this.rotation[2] -= 360;
+
+		/* Record frame */
+		this.addFrame("rotate");
 	}
 
 	public void zoom(double scale) {
 		//System.out.println("Drawer.zoom - " + scale);
-
 		this.location[2] *= scale;
+		this.addFrame("zoom");
 	}
 
 	public void quit() {
@@ -546,7 +550,7 @@ public class Drawer extends WindowAdapter implements
 	 */
 	public void actionPerformed(ActionEvent e) {
 		this.main.tick();
-		this.draw(plants);
+		this.addFrame("tick");
 	}
 
 	/*
@@ -569,6 +573,8 @@ public class Drawer extends WindowAdapter implements
 		System.out.format("Drawer.stateChanged - %s=%d - %d,%d,%d,%d\n",
 			name, value,
 			Main.water, Main.sun, Main.nutrition, Main.mutation);
+
+		this.addFrame("state");
 	}
 
 	/*
[…]

### 2014-03-28 src/plant_project/Main.java

@@ -112,12 +112,22 @@ public class Main {
 					time, time, time, time, time, time);
 				this.drawer.record(path);
 				break;
+			case 'f':
+				// record frame
+				this.drawer.addFrame("user");
+				break;
+			case 'x':
+				// kill all plants and reset
+				this.resetPlants();
+				break;
 			case ' ':
+				// play/pause
 				this.drawer.toggle();
+				break;
 			default:
 				// System.out.println("tick");
 				this.tick();
-				drawer.draw(plants);
+				this.drawer.addFrame("tick");
 				break;
 		}
 	}
@@ -158,42 +168,46 @@ public class Main {
 		System.out.println("tick - removed " + doomed);
 	}
 
-	/**
-	 * Main constructor
-	 */
-	public Main() {
-		this.plants = new ArrayList<Plant>();
-		this.drawer = new Drawer(this);
+	public void resetPlants() {
+		this.plants.clear();
 
 		/* Wikipedia example L-System */
-		//plants.add(new Plant(0, 0, new Gene(15,
+		//this.plants.add(new Plant(0, 0, new Gene(15,
 		//	new LRule[] { new LRule(1, 'X', "Gn[[X]sX]sG[sGX]nX"),
 		//	              new LRule(1, 'G', "GG") },
 		//	new Color(0.0f, 0.8f, 0.2f),
 		//	new Color(0.3f, 0.3f, 0.8f),
 		//	2, Plant.MAX_ITERATION)));
 
-		//plants.add(new Plant(0, 0, new Gene(15,
+		//this.plants.add(new Plant(0, 0, new Gene(15,
 		//	new LRule[] { new LRule(0.75, 'X', "GGGG[nX][sX][eX][wX][X]"),
 		//	              new LRule(0.75, 'G', "GG") },
 		//	new Color(0.0f, 0.8f, 0.2f),
 		//	new Color(0.3f, 0.3f, 0.8f),
 		//	2, Plant.MAX_ITERATION, 0.01f, 0.1f)));
 
-		plants.add(new Plant(0, 0, new Gene(15,
+		this.plants.add(new Plant(0, 0, new Gene(15,
 			new LRule[] { new LRule(0.75, 'X', "Gn[[X]sXL]eG[sGXF]X"),
 			              new LRule(0.75, 'G', "GG") },
 			new Color(0.0f, 0.8f, 0.2f),
[…]

### 2014-03-28 doc/paper.tex
Webpage, abstract

@@
 \usepackage[utf8]{inputenc}
+
+\newcommand{\onecol}[3]{
+  \begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{#1}
+    \vspace{0.0in}%??
+    \caption{#3}
+    \label{#2}
+  \end{figure}
+}
+
+\newcommand{\twocol}[3]{
+  \begin{figure*}
+    \centering
+    \includegraphics[width=5.5in]{#1}
+    \vspace{0.1in}%??
+    \caption{#3}
+    \label{#2}
+  \end{figure*}
+}
 
@@
 
-The project will focus on modeling natural looking plants using simulated
-environment parameters to control growth and evolution of different plant
-species. Genetic algorithms or other evolutionary techniques will be used to
-evolve plant suitable for growth in various environments.
+L-systems, or Lindenmayer systems, are often used for realistic looking
+artificial plant environments. While conceptually simple, L-systems have been
+extended with many features to create realistic plants models based on a
+description of the plant growth patterns. While significant progress has been
+made in providing robust frameworks for modeling artificial plants, the process
+of creating specific plants species is still very much a manual process that
+often requires a deep understanding of the tools used.
 
-Potential environmental conditions:
+In this project we will explore an alternate approach to plant species creation
+by using genetic algorithms to automatically evolve plants suitable to
+particular environments.
 
-\begin{itemize*}
-  \item Availability of water
-  \item Amount of sunlight
-  \item Soil conditions
-  \item Growth elevation
-\end{itemize*}
+By tuning environmental parameters for availability of water, the amount of
+sunlight, and the ambient soil conditions, artificial plant species are produced
+which resembling physical plants that grown in similar environments.
 
-Potential plant species parameters:
-
-\begin{itemize*}
-  \item Rate of water loss
-  \item Food and water storage capability
-  \item Total mass/volume
-  \item Crown height
-  \item Leaf area
-  \item Root structures
-\end{itemize*}
+In this approach to plant generation, the underlying L-system production rules
+are evolved based on random mutation combined with a fitness function to
+determine plant survival rates. In addition to the L-systems, several additional
+parameters are introduced to control plant growth and simulation at a more
+abstract level.
 
@@
 
-\section{Introduction}
-
-Karl Sims introduced Panspermia in 1990.
-
-LSystem
+\section{Software Architecture}
 
@@
 our project. The two frameworks that were the best candidates was
-OpenAlea\cite{openalea} and Algorithmic Botany\cite{abotany}.
-
-\section{Software Architecture}
+OpenAlea\cite{openalea} and Algorithmic Botany\cite{abotany}. Eventually we
+decided to implement the software from scratch.
 
@@
 architecture can be divided into two main categories: plant simulation using
-L-Systems and the graphical user interface.
+L-systems and the graphical user interface.
 
@@
 
-The L-Systems are implemented in the LSystem and LRule classes, which are
+The L-systems are implemented in the LSystem and LRule classes, which are
 responsible for storing the production rules associated with each plant and
@@
 Gene and Plant classes store representations of each living plant. The Gene
-class represents the plant genotype as a combination of the L-System production
+class represents the plant genotype as a combination of the L-system production
 rules along with additional information such as stem and leaf size, stem and
@@
 and contains it's X/Y position and orientation inside the world, along with the
-current age and a L-System string (L-String) representing the plants current
+current age and a L-system string (L-string) representing the plants current
 shape. The Plant class is also responsible for calculating each individual
@@
 
-\includegraphics[width=\columnwidth]{images/architecture.png}
+\onecol{images/architecture.png}{arch}{Conceptual software architecture}
 
@@
 Each plant in our environment contains a string of symbol (which we call an
-"L-String") representing the plants current shape and size. Each symbol can
+"L-string") representing the plants current shape and size. Each symbol can
 represent a stem segment, a flower, a leaf, etc. The plants genotype contains a
@@
 When the plant grows, the production rules are used to replace symbols in the
-plants current L-String with sequences of tokens to produces a larger and more
+plants current L-string with sequences of tokens to produces a larger and more
 complex plant.
 
-\subsection{L-System Genotypes}
+\onecol{images/screenshot.png}{screen}{Example plant simulation}
 
-We implemented a modified variant of L-Systems where a replacement probability
+\subsection{L-system Genotypes}
+
+We implemented a modified variant of L-systems where a replacement probability
 is included in addition to the matching and replacement parts of each production
@@
   \item[Growth rules] \hfill \\
-    The L-System productions used to control the plants growth are changed. The
+    The L-system productions used to control the plants growth are changed. The
     plant may become either more or less complex.
@@
 
-Overall, we have shown that artificial plants based on L-Systems can be evolved
-in realistic ways based on abstract environmental parameters. Using L-Strings to
+Overall, we have shown that artificial plants based on L-systems can be evolved
+in realistic ways based on abstract environmental parameters. Using L-strings to
 represent plants results in both realistic looking shapes and provides the

### 2014-03-28 doc/paper.tex
Start proofreading

@@
 
-L-systems, or Lindenmayer systems, are often used for realistic looking
+L-systems, or Lindenmayer systems, are often used to create realistic looking
 artificial plant environments. While conceptually simple, L-systems have been
 extended with many features to create realistic plants models based on a
-description of the plant growth patterns. While significant progress has been
-made in providing robust frameworks for modeling artificial plants, the process
-of creating specific plants species is still very much a manual process that
-often requires a deep understanding of the tools used.
+description of the plant growth patterns. Significant progress has been made in
+providing robust frameworks for modeling artificial plants, but the process of
+creating specific plants species is still very much a manual process that often
+requires a deep understanding of the tools used.
 
 In this project we will explore an alternate approach to plant species creation
-by using genetic algorithms to automatically evolve plants suitable to
-particular environments.
+using genetic algorithms to automatically evolve plants suitable to particular
+environments.
 
-By tuning environmental parameters for availability of water, the amount of
+By tuning environmental parameters for the availability of water, the amount of
 sunlight, and the ambient soil conditions, artificial plant species are produced
-which resembling physical plants that grown in similar environments.
+which resembling physical plants that may grown in similar environments.
 
-In this approach to plant generation, the underlying L-system production rules
-are evolved based on random mutation combined with a fitness function to
-determine plant survival rates. In addition to the L-systems, several additional
-parameters are introduced to control plant growth and simulation at a more
+With this approach to plant species generation, the underlying L-system
+production rules are evolved based on random mutation combined with a fitness
+function to determine plant survival rates. In addition to the L-systems,
+several additional parameters are introduced to control plant growth and
+simulation at a more
 abstract level.
@@
 
-We closely examined different existing frameworks that we considered using for
-our project. The two frameworks that were the best candidates was
-OpenAlea\cite{openalea} and Algorithmic Botany\cite{abotany}. Eventually we
-decided to implement the software from scratch.
+We closely examined several existing frameworks that we considered using for
+this project. The two frameworks that were the best candidates were Algorithmic
+Botany\cite{abotany} and OpenAlea\cite{openalea}.
 
-The software was implemented using Java and OpenGL for drawing. The top-level
-architecture can be divided into two main categories: plant simulation using
-L-systems and the graphical user interface.
+The Algorithmic Botany suit consists of the VLAB software which can be used on
+Mac OS and Linux, the L-studio software for Microsoft Windows, and an easy to
+use TreeSketch application for iPad. We primarily experimented with using VLAB
+under but found that a significant amount of effort was required to install and
+interact with the tools and we estimated that there would be a large learning
+curve before we would proficient enough to modify the software for our needs.
+
+OpenAlea is implemented as a set of Python libraries utilizing native code for
+certain functions. As with Algorithmic Botany, installation and use of the tools
+was challenging. OpenAlea appears to be most targeted towards plant researchers
+who wish to manually model specific plants.
+
+As L-systems are conceptually very simple, we decided to create a separate
+implementation of the algorithms in order to avoid risks associated with reusing
+existing software which we have little experience with. Our software was
+implemented using Java and OpenGL for drawing. The top-level architecture can be
+divided into two main categories: plant simulation using L-systems and the
+graphical user interface.
 
@@
 
-The L-systems are implemented in the LSystem and LRule classes, which are
-responsible for storing the production rules associated with each plant and
+Our L-systems algorithms are implemented in the LSystem and LRule classes, which
+are responsible for storing the production rules associated with each plant and
 iterating the plants phenotype as the plant grows throughout the simulation. The
@@
 shape. The Plant class is also responsible for calculating each individual
-plants fitness values at the end of each simulation tick.
+plants fitness values during each simulation tick.
 
@@
 Each plant in our environment contains a string of symbol (which we call an
-"L-string") representing the plants current shape and size. Each symbol can
+``L-string'') representing the plants current shape and size. Each symbol can
 represent a stem segment, a flower, a leaf, etc. The plants genotype contains a
@@
 
-An example of a plant at first generation:
+Example of a plant at the first generation:
 
@@
 each of the currently living plant objects becomes older. We wanted to have a
-similar approach to have plants grow in the real world. Consequently, during the
+similar approach to how plants grow in the real world. Consequently, during the
 first 4 ticks the plant grows as a seed under ground. On the fifth tick it
@@
 be extreme in any ways. Gene mutation occurs by choosing one of the production
-rules and then mutation the replacement side of the rule. A symbol is randomly
+rules and then mutating the replacement side of the rule. A symbol is randomly
 chosen from the replacement to be mutated.
 
-The mutation can either be that the symbols is simply deleted, or that is then
-replaced with a sequences from a predefined set of possible mutations.
+The mutation can either result in a symbol being deleted or begin replaced with
+a sequences from a predefined set of possible mutations.
 
@@
 groups, however these divisions are just for convenience and are not used by the
-mutation algorithm.
+mutation algorithm:
 
@@
 
-For the purpose of fitness calculations, a ``genetic pressure'' version of each
-of these parameters is calculated and used in the fitness function.
+For the purpose of fitness calculations, a ``genetic pressure'' is calculated
+for of each of these parameters and used in the fitness function.
 \[pressure = 1 - (parameter / 100)\]
@@
 
-When calculating each plants fitness several plant parameters are used:
+While evaluating the fitness function, several plant parameters are used and
+derived parameters are created to simplify the fitness function itself:
 
@@
   \item[Stem ratio]    \hfill \\
-    Normalized stem width, $0 .. 1$
+    Normalized stem width, 0 = narrow stems, 1 = wide stems
   \item[Leaf ratio]    \hfill \\
-    Normalized leaf size, $0 .. 1$
+    Normalized leaf size, 0 = small leaves, 1 = wide leaves
   \item[Stem count]    \hfill \\
@@
     The total storage space available to the plant. We do not calculate an
-    actual volume but instead just use a combination of the stem width from the
+    actual volume but instead use a combination of the stem width from the
     plants genes and the number of stem segments.
@@
   \item[Surface area]  \hfill \\
-    The total surface area available in the leaves for photosynthesis. Like the
-    storage space, this is a combination of the leaf size and the number of
+    The total surface area available in the leaves for photosynthesis. Similar
+    to storage space, this is a combination of the leaf size and the number of
     leaves.
@@
 
-The plant fitness function is divided by water, sun, an nutrient fitness. Each
-of these is calculated separately and the total fitness is the sum. We
-implemented the fitness equations such that if all the environmental parameters
-are equal, then all the fitness values will also be equal and the plants will
-evolve randomly. In a more comprehensive system, some plants could be considered
-more or less fit in any environment including an environment with equal
-environmental parameters.
+The plant fitness function is divided into three parts corresponding to the
+water, sun, an nutrient levels. Each of these is calculated separately and the
+total fitness is summed afterwards. We implemented the fitness equations such
+that if all the environmental parameters are equal, then all the fitness values
+will also be equal and the plants will evolve randomly. In a more comprehensive
+system, some plants could be considered more or less fit in any environment
+including an environment with equal environmental parameters.
 
@@
   \item More stem segments will result in a taller plant which can college more
-    light. We do not calculate actual heights, but assume that stem segments
-    improve light gathering.
+    light. We do not calculate actual heights, but assume that additional stem
+    segments improve light gathering.
 \end{itemize}
@@
 taking into consideration environmental information related to the neighboring
-plants in the environment. This would use more advanced lighting and collision
-detection to calculate sunlight based on shadows in the environment.
+plants. This would require more advanced lighting and collision detection to
+calculate sunlight collection and photosynthesis based on shadows in the
+environment.
 
@@
 In our system, a leafy plant may always win over other plants in a shady
-simulation environment even though such as large pant may not be support in a
+simulation environment even though such as large plant may not be support in a
 physical environment to do water or soil limitations.
 
-Some difficulty was encountered evolving plants to appear realistic, and many
-limitations are inherit which are difficult to simulation in a computer graphics
-setting. For example, two physical plants which share a remarkable resemblances
-my be suited to different environments due to small evolutionary changes. Waxy
-leaves, for example, could result in a leafy plant well suited to a dry
-environment.
+Some difficulty was encountered evolving plants that appear realistic; some
+limitations may be inherit to genetic algorithms which rely on graphical
+simulations. For example, two physical plants may share a remarkable resemblance
+but may be suited to different environments due to small evolutionary changes.
+Waxy leaves, for example, could result in a leafy plant well suited to a dry
+environment. Additionally, flowering plants often evolve to attract insects and
+other animals which act as pollinators and seed dispersers. Without complex
+organisms interacting with the plants it may be difficult to evolve certain
+features such as fruits and flowers, without predefining certain
+characteristics.
 
